Efficient online Sequential Monte Carlo smoother for partially observed stochastic differential equations

### 2017-02-12 ParisEM.tex
minor changes

Let $h$ be a function defined on $\{1,\ldots,N\}$,
\begin{align*}
\mathbb{E}\left[h(J^{i,j}_k)\middle| \mathcal{F}_k^N\right] & = \sum_{m\ge 1}\mathbb{E}\left[h(I_m)\1_{\tau=m}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}h(m)\left(\prod_{\ell=1}^{m-1}\mathbb{E}\left[\1_{(\mathcal{A}^k_{\ell})^c}\middle| \mathcal{F}_k^N\right]\right)\mathbb{E}\left[h(I_m)\1_{\mathcal{A}^k_{m}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}\left(\prod_{\ell=1}^{m-1}\mathbb{E}\left[1-\frac{\widehat{q}^{\Delta t_k}(\xi_{k-1}^{I_\ell},\xi_k^{i},\Lambda_{\ell})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\right)\\
&\hspace{5cm}\times\mathbb{E}\left[h(I_m)\frac{\widehat{q}^{\Delta t_k}(\xi_{k-1}^{I_m},\xi_k^{i},\Lambda_{\ell})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}\left(\mathbb{E}\left[1-\frac{q^{\Delta t_k}_{\theta}(\xi_{k-1}^{I_1},\xi_k^{1})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\right)^{m-1}\mathbb{E}\left[h(I_m)\frac{q^{\Delta t_k}_{\theta}(\xi_{k-1}^{I_m},\xi_k^{1})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \mathbb{E}\left[h(I_1)q^{\Delta t_k}_{\theta}(\xi_{k-1}^{I_1},\xi_k^{i})\middle| \mathcal{F}_k^N\right]/\mathbb{E}\left[q^{\Delta t_k}_{\theta}(\xi_{k-1}^{I_m},\xi_k^{i})\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{\ell=1}^N \frac{h(\ell)\omega_{k-1}^{\ell}q^{\Delta t_k}_{\theta}(\xi_{k-1}^{\ell},\xi_k^{i})}{\sum_{m=1}^N\omega_{k-1}^{m}q^{\Delta t_k}_{\theta}(\xi_{k-1}^{m},\xi_k^{i})}\eqsp,\\
&= \sum_{\ell=1}^N \Lambda_{k-1}^N(i,\ell)h(\ell) \eqsp,
\end{align*}
which concludes the proof.

@@

The independence is ensured by the mechanism of SMC methods.\\
By \eqref{eq:random:weight},

@@

\mathbb{E}\left[\tilde{\omega}^i_{k+1}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right] = \mathbb{E}\left[\frac{M^{-1}\sum_{m=1}^M \widehat{\qk}(\xi_{k}^{I^{i}_{k+1}}, \xi^{i}_{k+1};\zeta^m_{k+1})g_{k+1}(\xi^{i}_{k+1})}{\vartheta_{k+1}(\xi^{I^{i}_{k+1}}_{k}) p_{k+1}(\xi_{k}^{I^{i}_{k+1}},\xi^{i}_{k+1})}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right]\eqsp.

@@

Moreover, The p.d.f. of the couple $(\xi_{k+1}^i,I_{k+1}^i)\vert \mathcal{F}_k^N$ is given by
p(x,j) &= \frac{\omega_k^j\vartheta_{k+1}(\xi_k^j)p_k(\xi_k^j,x)}{\sum_{l=1}^N\int \omega_k^l\vartheta_{k+1}(\xi_k^l)p_k(\xi_k^l,y)\rmd y}=\frac{\omega_k^j\vartheta_{k+1}(\xi_k^j)p_k(\xi_k^j,x)}{\phi_k^N[\vartheta_{k+1}]}

@@

\mathbb{E}\left[\tilde{\omega}^i_{k+1}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right]&= \left(\phi^N_{k}[\vartheta_{k+1}]\right)^{-1} \sum_{j=1}^N\omega_k^j \int \vartheta_{k+1}(\xi^{j}_{k})\frac{\qk(\xi_{k}^{j},x) g_{k+1}(x)}{\vartheta_{k+1}(\xi^{j}_{k}) p_{k}(\xi_{k}^{j},x)}\\

@@

&~~~~\times\sum_{\ell=1}^N \omega_k^\ell\left[\int \frac{ \sum_{j=1}^N \omega_k^j\qk(\xi_k^j,x) }{ \sum_{\ell'=1}^N\omega_k^{\ell'}\qk(\xi_{k}^{\ell'},x) } g_{k+1}(x)\qk (\xi_{k}^{\ell},x)\left(\tau^{\ell}_k + h_{k}(\xi_{k}^{\ell},x)\right) \rmd x \right]\\ 

>>>>

Let $h$ be a function defined on $\{1,\ldots,N\}$,
\begin{align*}
\mathbb{E}\left[h(J^{i,j}_k)\middle| \mathcal{F}_k^N\right] & = \sum_{m\ge 1}\mathbb{E}\left[h(I_m)\1_{\tau=m}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}h(m)\left(\prod_{\ell=1}^{m-1}\mathbb{E}\left[\1_{(\mathcal{A}^k_{\ell})^c}\middle| \mathcal{F}_k^N\right]\right)\mathbb{E}\left[h(I_m)\1_{\mathcal{A}^k_{m}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}\left(\prod_{\ell=1}^{m-1}\mathbb{E}\left[1-\frac{\widehat{q_k}(\xi_{k-1}^{I_\ell},\xi_k^{i};\zeta_{\ell})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\right)\\
&\hspace{5cm}\times\mathbb{E}\left[h(I_m)\frac{\widehat{q_k}(\xi_{k-1}^{I_m},\xi_k^{i};\zeta_{\ell})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{m\ge 1}\left(\mathbb{E}\left[1-\frac{q_k(\xi_{k-1}^{I_1},\xi_k^{1})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\right)^{m-1}\mathbb{E}\left[h(I_m)\frac{q_k(\xi_{k-1}^{I_m},\xi_k^{1})}{\sigma_{+}}\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \mathbb{E}\left[h(I_1)q_k(\xi_{k-1}^{I_1},\xi_k^{i})\middle| \mathcal{F}_k^N\right]/\mathbb{E}\left[q_k(\xi_{k-1}^{I_m},\xi_k^{i})\middle| \mathcal{F}_k^N\right]\eqsp,\\
& = \sum_{\ell=1}^N \frac{h(\ell)\omega_{k-1}^{\ell}q_k(\xi_{k-1}^{\ell},\xi_k^{i})}{\sum_{m=1}^N\omega_{k-1}^{m}q_k(\xi_{k-1}^{m},\xi_k^{i})}\eqsp,\\
&= \sum_{\ell=1}^N \Lambda_{k-1}^N(i,\ell)h(\ell) \eqsp,
\end{align*}
which concludes the proof.

@@

The independence is ensured by the mechanism of SMC methods. By \eqref{eq:random:weight},

@@

\mathbb{E}\left[\widehat{\omega}^i_{k+1}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right] = \mathbb{E}\left[\frac{M^{-1}\sum_{m=1}^M \widehat{\qk}(\xi_{k}^{I^{i}_{k+1}}, \xi^{i}_{k+1};\zeta^m_{k+1})g_{k+1}(\xi^{i}_{k+1})}{\vartheta_{k+1}(\xi^{I^{i}_{k+1}}_{k}) p_{k+1}(\xi_{k}^{I^{i}_{k+1}},\xi^{i}_{k+1})}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right]\eqsp.

@@

Moreover, conditionally to $\mathcal{F}_k^N$, the probability density function of $(\xi_{k+1}^i,I_{k+1}^i)$ is given by
(x,j) \mapsto \frac{\omega_k^j\vartheta_{k+1}(\xi_k^j)p_k(\xi_k^j,x)}{\Omega_k\phi_k^N[\vartheta_{k+1}]}\eqsp.

@@

\mathbb{E}\left[\widehat{\omega}^i_{k+1}\tau^{i}_{k+1}\middle| \mathcal{F}_k^{N}\right]&= \left(\phi^N_{k}[\vartheta_{k+1}]\right)^{-1} \sum_{j=1}^N\frac{\omega_k^j}{\Omega_k} \int \vartheta_{k+1}(\xi^{j}_{k})\frac{\qk(\xi_{k}^{j},x) g_{k+1}(x)}{\vartheta_{k+1}(\xi^{j}_{k}) p_{k}(\xi_{k}^{j},x)}\\

@@

&~~~~\times\sum_{\ell=1}^N \frac{\omega_k^\ell}{\Omega_k}\left[\int \frac{ \sum_{j=1}^N \omega_k^j\qk(\xi_k^j,x) }{ \sum_{\ell'=1}^N\omega_k^{\ell'}\qk(\xi_{k}^{\ell'},x) } g_{k+1}(x)\qk (\xi_{k}^{\ell},x)\left(\tau^{\ell}_k + h_{k}(\xi_{k}^{\ell},x)\right) \rmd x \right]\\ 
